Distribution-oblivious Programming with Dish

### 2019-09-21 pldi/draft.tex
Update intro

\section{Background}

@@

\subsection{Unix Pipelines, Informally}

Shell scripting

@@

@@

\heading{Generation}
\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
 cat * | wc -l > out.txt                                          (($$p_2$$))
\end{lstlisting}

\heading{Identifiers}
\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
 cat /proc/ >                                                   (($$p_2$$))
\end{lstlisting}

\heading{Stream Manipulation}
Explicit Redirection
Implicit splitting and merging 
This can be done through the file-system operators

\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
 cat /proc/ >                                                   (($$p_2$$))
\end{lstlisting}

\heading{Grouping}
Composition

@@

\subsection{Comnands}

@@

\subsection{The Shell Language}

>>>>

\section{Background}
\label{bg}

@@

\subsection{Unix Pipelines, Informally}
\label{bg:pipelines}

Pipelines are a mechanism for program composition.
Individual programs, specified as commands, conform to a common but general interface. 
Broadly, this interface commands allows commands to do a combination of the following:
  accept input from their input stream (stdin),
  generate output to their output stream (stdout)
Input and output streams are standard text streams---lines of text separated by the newline character.
Leveraging this interface, pipelines chain together commands by their standard streams, such that the output stream of one command (stdout) is passed directly as input (stdin) to the next one.
% This concept gave rise to the toolbox philosophy: 
%   programs that do one thing and do it well, that work together, and that handle text streams as a universal interface.

@@

The standard shell syntax for anonymous pipes is to list multiple commands, separated by vertical bars ("pipes" in common Unix verbiage).
that are not interpreted by the shell, but are local to each command.

Additional details augment this simple interface.
Commands can be specified with options arguments

@@

\heading{Generation}
A central concept in pipelines---and Unix, in general---is the file system, which allows interfacing with persistent data.
There are many commands that take as input a file identifier 
Reading a file or a set of files generates a stream; 
  similarly, streams can be trivially redirected to the file system---by appending with the file identifier.

Expansion---the key is that expansion and evaluation are handled by the shell at an earlier stage 
The most powerful---and interesting, for \ttt{dish}---type of expansion is subshell expansion, that allows replacing entire input streams

\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
 cat * | wc -l > out.txt                 (($$p_3$$))
\end{lstlisting}

\heading{Identifiers}
The file-system is a common abstraction layer used for many other operations than just serving files (this will be become clearer when talking about file-system commands in \S\ref{bg:cmd}).
as such, different identifiers many not resolve to actual files, but rather in-memory or OS-internal structures.

Unix already specifies a few classes of files:

Here are a few special files:
  \ttt{proc} called

Technically, anonymous pipes show up as file identifiers in a hidden filesystem.
Similarly, open files and other ephemeral identifiers

\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
 cat /proc/ >                            (($$p_4$$))
\end{lstlisting}

\heading{Stream Manipulation}
A small but powerful set of primitives are concerned with stream manipulation.
First, as described above, several commands can be redirected
Second,

\ttt{xargs}
\ttt{tee}

In combination, these constructs are powerful enough to introduce circular streams. % or stream cycles?

Explicit Redirection
Implicit splitting and merging 
This can be done through the file-system operators

\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
 cat /proc/ >                            (($$p_5$$))
\end{lstlisting}

\heading{Composition}
Composition

@@

\subsection{Comnands}
\label{bg:cmd}

@@

\subsection{The Shell Language}
\label{bg:shell}
